---
title: "Homework A"
subtitle: "Intro to R and RStudio"
author: "YOUR NAME HERE"
format:
  html:
    toc: true
    code-overflow: wrap
    code-fold: true
    embed-resources: true
execute: 
  message: FALSE
  warning: FALSE
---

# In-lab Examples 

bold

italics

bold and italics

a list

link

image

---
R chunks

create an object

inline code

commenting code

# Questions 

1.  Tell me your name in bold lettering. Is there anything we should know about your name and how to address you? (ie. I prefer to be called Liz instead of Elizabeth, my pronouns are he/his/him, etc.)
 

2. Make a list of a few of your academic interests. 


3. Include a link to delicious restaurant in Sendai. 


4.  Include a picture. This could be a picture of you or a picture of something you find funny or interesting.  


5.  Create an object called `height`. Assign your height in m to this object (you can also use a made up number here if you like). Use R to convert the height to inches. Be sure to comment your code to explain what you are doing. 

```{r}

```

6. Create an object called `c`. Decide on a number between 1 and 10 that denotes your comfort level with taking this R workshop (10 being that you are so, so excited and 1 being that you are so nervous you almost threw up this morning just thinking about it). Assign this value to `c`. 

```{r}

```

7.  Write a few sentences about why you choose this number. Be sure to print out this value of `c` somewhere in your paragraph as inline code.


8. Are there any topics or subjects that you really hope are covered in this workshop? What are you hoping to learn that will be most useful to you and your work?


9.  (Optional) Use this space below to tell me anything about yourself that you think Easton and Emily should know in order to teach you effectively. You can also use this space below to ask any questions or state any concerns.



