---
title: "ANOVA_Notes_Blank"
format: html
editor: visual
---


```{r}
library(tidyverse)
```


#ANOVA (Analysis of variance) with cuckoos data set

An analysis of variance (ANOVA) is a statistical test used to determine whether a continuous and a categorical variable differ. The categorical variable should have three or more groups, or levels. 


```{r}
library(DAAG)
data("cuckoos")
```

Research question: Does egg length differ between the six species?

Null hypothesis: 

$$
H_0: \mu_{\text{hedgesparrow}} = \mu_{\text{meadowpipit}} = \mu_{\text{piedwagtail}} = \mu_{\text{robin}} = \mu_{\text{treepipit}} = \mu_{\text{wren}}
$$

Alternative hypothesis:

$$
H_a: \text{At least one mean is different } \mu_{\text{species}} \neq \mu_{\text{another species}}
$$

##Visualizing the data

```{r}
#box plot

```


##ANOVA test assumptions

-   Independence: The data collected is from a representative and randomly selected proportion of the total population.

-   Normality: The data should be normally distributed

-   Equality of variances: The variances should be equal across the groups (homoscedasticity).

-  Outliers: There should be no significant outliers in different groups.

```{r}
```


##Running the anova

```{r}

```


#ANOVA with the penguins data set

```{r}
library(palmerpenguins)
```

Research question: does flipper length differ between species?

##Visualizing the data

```{r}

```

##Checking assumptions

```{r}

```


##Running the ANOVA

```{r}

```

