---
title: "ANOVA_Notes"
format: html
editor: visual
---

```{r}
library(tidyverse)
```


#ANOVA (Analysis of variance) with cuckoos data set

An analysis of variance (ANOVA) is a statistical test used to determine whether a continuous and a categorical variable differ. The categorical variable should have three or more groups, or levels. 


```{r}
library(DAAG)
data("cuckoos")
```

Research question: Does egg length differ between the six species?

Null hypothesis: 

$$
H_0: \mu_{\text{hedgesparrow}} = \mu_{\text{meadowpipit}} = \mu_{\text{piedwagtail}} = \mu_{\text{robin}} = \mu_{\text{treepipit}} = \mu_{\text{wren}}
$$

Alternative hypothesis:

$$
H_a: \text{At least one mean is different } \mu_{\text{species}} \neq \mu_{\text{another species}}
$$

##Visualizing the data

```{r}
#box plot
ggplot(data = cuckoos, aes(x = species, y = length))+
  geom_boxplot()+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```


##ANOVA test assumptions

-   Independence: The data collected is from a representative and randomly selected proportion of the total population.

-   Normality: The data should be normally distributed

-   Equality of variances: The variances should be equal across the groups (homoscedasticity).

-  Outliers: There should be no significant outliers in different groups.

```{r}
ggplot(data = cuckoos,aes(x = length))+
  geom_histogram()

cuckoos %>% 
  group_by(species) %>% 
  summarise(var(length))
```




##Running the anova

```{r}
model1 <- aov(length ~ species, data = cuckoos)
model2 <- lm(length ~ species, data = cuckoos)
tukey <- TukeyHSD(model1)#make sure your object is called "tukey" for the code below. 
```


```{r}
#| echo: true

#Extracting significant pairs
tukey_df <- as.data.frame(tukey$species)
names(tukey_df)[names(tukey_df) == "p adj"] <- "p.adj"
sig_pairs <- subset(tukey_df,p.adj < 0.05)

# Create list of comparisons from significant pairs
comparisons <- strsplit(row.names(sig_pairs), "-")
#Creating the plot
library(ggsignif)
ggplot(cuckoos, aes(x = species, y = length)) +
  geom_boxplot() +
  geom_signif(comparisons = comparisons,
  annotations = ifelse(sig_pairs$p.adj < 0.001,   "***",                               ifelse(sig_pairs$p.adj < 0.01, "**",               ifelse(sig_pairs$p.adj < 0.05, "*", ""))),
y_position = seq(25, 35, length.out = nrow(sig_pairs)), #changing location of the bars
tip_length = 0.02) +
theme_bw()+
theme(axis.text.x = element_text(angle = 30, 
    hjus = 1, size = 18),
    axis.text.y = element_text(size = 18),          
    axis.title.x = element_text(size = 15),         
    axis.title.y = element_text(size = 20)) +
  labs(x = "Host species", y = "Egg length (mm)")
```



#ANOVA with the penguins data set

```{r}
library(palmerpenguins)
data("penguins")
View(penguins)
```

Research question: does flipper length differ between species?

##Visualizing the data

```{r}
ggplot(data = penguins, aes(x = species, y = flipper_length_mm))+
  geom_boxplot()



```

##Checking assumptions

```{r}

ggplot(data = penguins, aes(x = flipper_length_mm))+
  geom_histogram()+
  facet_wrap(~species)


penguins %>% 
  group_by(species) %>% 
  summarise(var(flipper_length_mm,  na.rm = TRUE))

```

##Running the ANOVA

```{r}
#using aov
mod1 <- aov(flipper_length_mm ~ species, data = penguins)
summary(mod1)#getting output of the model

tukey <- TukeyHSD(mod1)
tukey

#using lm
mod2 <- lm(flipper_length_mm ~ species, data = penguins)
summary(mod2)



```
